---
title: "LAJC Wage Theft Data"
author: 'Jon Kropko -- Code for Charlottesville'
date: "10/1/2019"
output: 
     html_document:
          toc: TRUE
          toc_depth: 4
          toc_float: true
          toc_collapsed: true
          theme: journal
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r lib, warning=FALSE, message=FALSE}
library(knitr)
library(lubridate)
library(tidyverse)
```

## Raw data
First I load the raw data.
```{r loaddata}
wagetheft <- read_csv("lajc_wage_claim.csv")
```

## Goal 2: A table showing a comprehensive overview of the reasons that DOLI has rejected cases over the last five years.
For each reason, we’d like to know the number of cases dismissed on this basis (for example: "47 cases were dismissed because the worker worked some overtime"). (Some cases have more than one reason for dismissal.)

This unfortunately may require a human being to look over the various reasons for dismissal listed in the “Other” columns (columns 21 and 31) and determine how to code each one. This may require a little independent judgment in some cases. For example, DOLI will reject a case if the worker is seeking over $15,000. One DOLI investigator may note this as "$15,000 limit," while another may write "exceeds maximum claim amount." 

The reasons for dismissal that LAJC is most interested are these:

  * The business is closed / out of business (this is different from “business bankrupt”)
  
  * The worker had a written employment agreement
  
  * The worker worked overtime in at least one week
  
  * The work was performed outside of Virginia
  
  * The worker is seeking tips or was a tipped employee
  
  * The claim amount is over $15,000  
  
  * The worker hired an attorney (this is different from “the worker filed a lawsuit”)
  
  * The worker did not demand the wages from the employer before filing a complaint with DOLI
  
  * The worker was an independent agent/subcontractor/self-employed or DOLI was otherwise unable to establish an employer-employee relationship
  
  * Errors with the claim form (e.g. the form was incomplete, or the worker didn’t provide an exact claim amount, or the worker signed in pencil rather than in pen, etc.)
  
  * The claim seeks minimum wages
  
  * But please track the other reasons as well! (E.g., complainant can’t be found / worker dropped or withdrew their claim / worker got paid / 2-year statute of limitations / insufficient documentation to proceed, etc.)
  
## Data cleaning
I first keep only the variables I need for this work:
```{r clean}
wagetheft <- wagetheft %>%
  select(`CLAIM NO`, Invalid:`CLAIM INVAL OTHER DESCRIPTION`)
```
I suspect that only one reason is given for any judgement that a claim is fund to be invalid. But to confirm this, I check whether there is ever more than one non-missing cell for the "Fringe Benefits", "Independent Agent", "Subcontractor", "False Claim", and "Other" variables.
```{r nonmissing}
check <- !is.na(select(wagetheft, `Fringe Benefits`:Other))
table(rowSums(check))
```
This check shows that `sum(rowSums(check))` observations have exactly 1 stated reason for a judgement of an invalid claim, and the other `sum(1-rowSums(check))` observations are not marked invalid. No case has more than one reason listed for being invalid.

## Bringing the categories into one column
To make a publishable-quality table, I need to create one variable that combines all of these reasons:
```{r onevariable}
wagetheft <- wagetheft %>%
  mutate(invalid_reason = as.numeric(!is.na(`Fringe Benefits`)) +
           2*as.numeric(!is.na(`Independent Agent`)) + 3*as.numeric(!is.na(Subcontractor)) +
           4*as.numeric(!is.na(`False Claim`)) + 5*as.numeric(!is.na(Other)),
         invalid_reason = factor(invalid_reason),
         invalid_reason = fct_recode(invalid_reason,
                                     "Not found invalid" = "0",
                                     "Fringe benefits" = "1",
                                     "Independent agent" = "2",
                                     "Subcontractor" = "3",
                                     "False claim" = "4",
                                     "Other" = "5"))
table(wagetheft$invalid_reason)
```
However, before we can move forward with a complete table, we need to delve into the reasons stated under "Other". These reasons are listed in "CLAIM INVAL OTHER DESCRIPTION":
```{r other}
kable(table(wagetheft$`CLAIM INVAL OTHER DESCRIPTION`))
```
We need to recode these reasons in the main `invalid_reason` variable. I write a function to do this efficiently:
```{r othercats}
other <- function(old, new){
  wagetheft <- wagetheft %>%
  mutate(invalid_reason = ifelse(`CLAIM INVAL OTHER DESCRIPTION`==old, new, invalid_reason))
  return(wagetheft)
}
```
I then apply this function to the reasons listed in the data:
```{r otherapply}
wagetheft <- wagetheft %>%
  mutate(`CLAIM INVAL OTHER DESCRIPTION` = fct_recode(`CLAIM INVAL OTHER DESCRIPTION`,)
         "" ="CANNOT ESTABLISH EMPLOYEE/EMPLOYER RELATIONSHIP",
         "" ="WORK PERFORMED OUTSIDE OF VA",
         "" ="EMPLOYMENT AGREEMENT",
         "" ="WORK PERFORMED OUTSIDE OF VIRGINA",
         "" ="WRITTEN EMPLOYMENT AGREEMENT",
         "" ="NO GROSS CLAIM AMOUNT",
         "" ="NO EMPLOYEE/EMPLOYER RELATIONSHIP",
         "" ="CORPORATE DIRECTOR/HIRED ATTY",
         "" ="HIRED LAWYER",
         "" ="WAGES NOT DEMANDED",
         "" ="INVOLVES OVERTIME",
         "" ="WORK PERFORMED OUTSIDE OF VIRGINIA",
         "" ="WORKED OUTSIDE OF VIRGINIA",
         "" ="WORKED ON A BASE",
         "" ="DIRECTOR, OFFICER, OWNER OR PARTNER",
         "" ="CITY EMPLOYEE",
         "" ="SELF EMPLOYED",
         "" ="CLAIMANT RECD FUNDS FROM EMPLOYER FOR VACATION TIME",
         "" ="SELF-EMPLOYED",
         "" ="OVERTIME/WRITTEN AGREEMENT",
         "" ="NO CLAIM PERIOD",
         "" ="TIPPED EMPLOYEE",
         "" ="COLLECTIVE BARGAINING AGREEMENT",
         "" ="COLLECTIVE BARGAINING AGREMENT",
         "" ="CLAIMANT DID NOT SIGN FORM",
         "" ="CLAIMANT DID NOT CORRECTLY FILL OUT FORM, NEEDS MORE INFO",
         "" ="WORK PERFORMED OUTSIDE VA; HIRED LAWYER",
         "" ="WORKED IN WASHINGTON, DC",
         "" ="CASE INVOLVES OVERTIME",
         "" ="CLAIMANT WAS EMPLOYEE OF PUBLIC SCHOOL SYSTEM",
         "" ="CLAIMANT NEVER DEMANDED WAGES",
         "" ="CLAIM NOT SIGNED AND CLAIMANT SEEKING OVERTIME PAY",
         "" ="CLAIMANT HIRED A LAWYER",
         "" ="WRITTEN EMPLOYMENT AGREEMENT NOTED",
         "" ="CLAIMANT HAS WRITTEN EMPLOYMENT AGREEMENT",
         "" ="CLAIMANT HAS NOT YET DEMANDED WAGES",
         "" ="OVERPAID",
         "" ="OVERPAYMENT",
         "" ="CLAIMANT INCLUDED WRITTEN EMPLOYMENT AGREEMENT",
         "" ="CLAIMANT INCLUDED WRITTEN EMPLOYMENT AGREEMENT.",
         "" ="CLAIMANT PERFORMED WORK IN DC",
         "" ="CLAIMANT IS SELF EMPLOYED",
         "" ="CLAIM PERIOD OVER 2 YEARS AGO AND CLAIMANT MARKED SELF EMPLOYED",
         "" ="BUSINESS PARTNER",
         "" ="WAGE ADVANCEMENT",
         "" ="CLAIMANT DID NOT ATTACH, BUT INDICATED THEY HAD A WRITTEN EMPLOYMENT AGREEMENT",
         "" ="CLAIMANT INDICATED SHE HAS A WRITTEN EMPLOYMENT AGREEMENT",
         "" ="CLAIMANT INCLUDED A WRITTEN EMPLOYMENT AGREEMENT",
         "" ="CLAIMANT INDICATED HE HAS A WRITTEN EMPLOYMENT AGREEMENT",
         "" ="CLAIMANT IS SEEKING OVER 1 MILLION DOLLARS",
         "" ="CONFLICTING DOCUMENTATION",
         "" ="NO WORK PERFORMED",
         "" ="FISCAL INTERMEDIARY",
         "" ="CLAIMANT IS A GOVERNMENT WORKER",
         "" ="CLAIMANT INDICATED SHE HAS A WRITTEN EMPLOYMENT AGREEMENT AND CLAIM OVER $15,000",
         "" ="CLAIMANT SEEKING MONEY EARNED AS A TIPPED EMPLOYEE.",
         "" ="CLAIMANT INDICATED THEY HAVE A WRITTEN EMPLOYMENT AGREEMENT",
         "" ="CLAIM FORM NOT SIGNED IN INK",
         "" ="CLAIMANT NOTED THEY HAVE A WRITTEN EMPLOYMENT AGREEMENT",
         "" ="CLAIMANT INDICATED THEY HAD A WRITTEN EMPLOYMENT AGREEMENT AND WAS SELF-EMPLOYED",
         "" ="CLAIMANT DID NOT GIVE A CLAIM AMOUNT",
         "" ="CLAIMANT DID NOT SPECIFY A CLAIM RANGE AND CLAIM AMOUNT",
         "" ="CLAIMANT HAS A WRITTEN EMPLOYMENT AGREEMENT",
         "" ="CLAIM AMOUNT IS OVER $15,000 AND CLAIM PERIOD IS OVER 2 YEARS OLD",
         "" ="CLAIM FORM WAS COPIED",
         "" ="OVERTIME",
         "" ="CLAIMANT IS SELF EMPLOYED, HAS A WRITTEN EMPLOYMENT AGREEMENT, WORKED OUTSIDE VA",
         "" ="CLAIMANT HAS WRITTEN EMPLOYMENT AGREEMENT AND WORKED ON MILITARY BASE",
         "" ="CLAIMANT INDICATING SHE HAS A WRITTEN EMPLOYMENT AGREEMENT",
         "" ="CLAIMANT PROVIDED WRITTEN EMPLOYMENT AGREEMENT",
         "" ="CLAIMANT FILED IN COURT",
         "" ="CLAIMANT IS A PUBLIC EMPLOYEE AND HAS A WRITTEN EMPLOYMENT AGREEMENT.",
         "" ="CLAIMANT IS SELF-EMPLOYED.",
         "" ="CLAIMANT INCLUDED A WRITTEN EMPLOYMENT AGREEMENT.",
         "" ="CLAIMANT INDICATED HAVING A WRITTEN EMPLOYMENT AGREEMENT.",
         "" ="WORK PERFORMED IN MD AND VA",
         "" ="CLAIM INVOLVES TIPS",
         "" ="WORKED IN MULTIPLE STATES",
         "" ="WORK PERFORMED MORE THAN 2 YEARS AGO",
         "" ="$15,000 CLAIM LIMIT",
         "" ="CLAIMANT HAS A WRITTEN EMPLOYMENT AGREEMENT.",
         "" ="WORK PERFORMED OUTSIDE OF VIRGINIA.",
         "" ="CLAIMANT WAS A CORPORATE DIRECTOR.",
         "" ="CLAIMANT IS SELF-EMPLOYED/HAD A WRITTEN AGREEMENT.",
         "" ="CLAIMANT INDICATED HE HAD A WRITTEN EMPLOYMENT AGREEMENT.",
         "" ="CLAIMANT INDICATED HE WAS SELF-EMPLOYED.",
         "" ="CLAIM RELATES TO OVERTIME.",
         "" ="WORK PERFORMED IN MULTIPLE STATES",
         "" ="CLAIMANT PERFORMED WORK ON MILITARY INSTALLATIONS.",
         "" ="CLAIMANT ALREADY FILED IN COURT.",
         "" ="CLAIMANT PERFORMED WORK OUT OF STATE",
         "" ="PERFORMED WORK OUTSIDE VIRGINIA.",
         "" ="CLAIMANT SEEKING MINIMUM WAGE.",
         "" ="CLAIMANT FILED IN COURT.",
         "" ="DISCRETIONARY BONUS",
         "" ="CLAIMANT SEEKING OT PAY.",
         "" ="CLAIMANT IS A DIRECTOR, HAS A WRITTEN AGREEMENT. AND HIRED A LAWYER.",
         "" ="CLAIMANT WAS SELF EMPLOYED.",
         "" ="CLAIMANT SEEKING REGULAR TIME AND OVERTIME.",
         "" ="CLAIMANT WAS SELF-EMPLOYED.",
         "" ="CLAIMANT PERFORMED THE WORK IN MARYLAND.",
         "" ="CLAIMANT INCLUDED A PARTIAL WRITTEN EMPLOYMENT AGREEMENT.",
         "" ="CLAIMANT IS A TIPPED EMPLOYEE.",
         "" ="CLAIMANT HAD A WRITTEN EMPLOYMENT AGREEMENT.",
         "" ="CLAIMANT SEEKING OVERTIME PAY.",
         "" ="CLAIMANT INDICATED HAVING A WRITTEN EMPLOYMENT AGREMEENT.",
         "" ="CLAIMANT PERFORMED WORK IN MANY STATES",
         "" ="CLAIMANT DID NOT INCLUDE A TIME PERIOD FOR UNPAID WAGES.",
         "" ="OVERTIME CLAIM",
         "" ="CLAIMANT FAILED TO PROVIDE A CLAIM PERIOD AND TOTAL.",
         "" ="CLAIMANT INDICATED SHE HAD A WRITTEN EMPLOYMENT AGREEMENT.",
         "" ="CLAIMANT PERFORMED WORK OUTSIDE VIRGINIA.",
         "" ="CLAIM IS OVER 2 YEARS OLD.",
         "" ="CLAIMANT DID NOT INCLUDE A TIME PERIOD OR CLAIM AMOUNT.",
         "" ="CLAIMANT DID NOT INCLUDE A TIME PERIOD.",
         "" ="CLAIMANT IS A GOVERNMENT EMPLOYEE.",
         "" ="CLAIM INVOLVES OVERTIME",
         "" ="CLAIMANT INDICATED HAVING A WRITTEN EMPLOMENT AGREEMENT.",
         "" ="CLAIM INVOLVED OVERTIME.",
         "" ="CLAIMANT HAS NOT DEMANDED WAGES",
         "" ="CLAIMANT PERFORMED WORK ON FEDERAL PROPERTY.",
         "" ="DID NOT DEMAND WAGES",
         "" ="OUTSIDE SALES",
         "" ="CLAIMANT INDICATED HE HAD A WRITTEN EMPLOYMENT AGREEMENT",
         "" ="INCLUDED WRITTEN EMPLOYMENT AGREEMENT",
         "" ="WORK WAS PERFORMED OUT OF THE STATE",
         "" ="CLAIMANT INDICATED THAT HE WAS SELF EMPLOYED",
         "" ="CLAIMANT INDICATED THEY HAD A WRITTEN EMPLOYMENT AGREEMENT.",
         "" ="CLAIMANT ALREAD FILED IN COURT AGAINST THE EMPLOYER",
         "" ="CLAIMANT IS SEEKING OVERTIME",
         "" ="CLAIMANT PERFORMED THE WORK IN NORTH CAROLINA.",
         "" ="PAY DATE HAS NOT YET PASSED",
         "" ="CLAIM INVOLVES TIPS AND OVERTIME.",
         "" ="CLAIMANT PERFORMED WORK ON A MILITARY BASE",
         "" ="CLAIM INVOLVES OVERTIME.",
         "" ="CLAIMANT DID NOT INDICATE THE TOTAL AMOUNT OF UNPAID WAGES.",
         "" ="CLAIMANT HAD A WRITTEN EMPLOYMENT AGREEMENT AND HIRED A LAWYER",
         "" ="CLAIMANT INDICATED THE FILED A COURT CASE FOR UNPAID WAGES.",
         "" ="CLAIMANT DID NOT SIGN THE CLAIM FORM.",
         "" ="CLAIMANT HAS A WRTTEN EMPLOYMENT AGREEMENT.",
         "" ="CLAIM INVOLVING OVERTIME",
         "" ="CLAIMANT WORKED FOR THIS BUSINESS AS A SELF-EMPLOYED PERSON.",
         "" ="CLAIMANT WORKED ON A MILITARY INSTALLATION",
         "" ="TOTAL GROSS AMOUNT OF UNPAID WAGES NOT INCLUDED",
         "" ="EXCEEDS LIMIT FOR ACCEPTANCE",
         "" ="WORK PERFORMED OUTSIDE VA",
         "" ="CLAIMANT HAS FILED IN LOUDOUN COUNTY COURT",
         "" ="FILED CASE WITH WISE GENERAL DISTRICT COURT",
         "" ="CLAIMANT INDICATED SHE HAS A WRITTEN EMPLOYMENT AGREEMENT.",
         "" ="CLAIMANT DID NOT DEMAND HIS UNPAID WAGES",
         "" ="CLAIMANT ALREADY FILED IN COURT",
         "" ="TOTAL GROSS AMOUNT OF UNPAID WAGES ARE NOT INCLUDED",
         "" ="TOTAL AMOUNT OF UNPAID WAGES NOT INCLUDED",
         "" ="CLAIMANT DID NOT DEMAND WAGES",
         "" ="CLAIMANT SEEKING MINIMUM WAGE + TIPS",
         "" ="CLAIMANT DID NOT UNCLUDE SIGNATURE",
         "" ="CLAIMANT INDICATED THEY HAD A WRITTEN EMPLOYMENT AGREEMENT",
         "" ="CLAIMANT INDICATED HE WAS A CORPORATE DIRECTOR, OWNER OR PARTNER  IN BUSINESS",
         "" ="CLAIMANT INDICATED SHE HAD A WRITTEN EMPLOYMENT AGREEMENT",
         "" ="CLAIMANT INDICATED HE WAS A CORPORATE DIRECTOR, OWNER OR PARTNER.",
         "" ="CLAIMANT DID NOT INCLUDE THE TOTAL GROSS AMOUNT OF UNPAID WAGES",
         "" ="CLAIMANT'S TOTAL GROSS AMOUNT EXCEEDS THE $15,000 LIMIT",
         "" ="CLAIMANT DID NOT DEMAND THEIR WAGES",
         "" ="MINIMUM WAGE",
         "" ="CLAIMANT PERFORMED WORK IN TUSCON, AZ.",
         "" ="CLAIMANT REQUESTED TIPS",
         "" ="MINIMUM WAGE CLAIM",
         "" ="CLAIMANT SEEKING OVERTIME PAY",
         "" ="CLAIMANT IS SEEKING OVERTIME PAY",
         "" ="CLAIMANT HAD A WRITTEN EMPLOYMENT AGREEMENT",
         "" ="CLAIMANT DID NOT INCLUDE TOTAL GROSS AMOUNT OF UNPAID WAGES",
         "" ="CLAIMANT SEEKING WAGES UNDER MINIMUM WAGE.",
         "" ="CLAIMANT DID NOT INCLUDE TOTAL AMOUNT OF UNPAID WAGES",
         "" ="CLAIMNANT'S SEEKING OVERTIME",
         "" ="WRITTEN EMPLOYMENT AGREEMENT INCLUDED",
         "" ="CLAIMANT HAS NOT BEEN PAID AT HIGHER RATE",
         "" ="VOIDED SALES & SALES NOT MADE BY CLAIMANT",
         "" ="COMPLAINANT DID NOT WORK HOURS CLAIMED.",
         "" ="CLAIMANT WAS PAID FOR 80 HOURS THAT HE DID NOT WORK WHILE HOSPITALIZED. CLAIM WA",
         "" ="EMPLOYER DOCUMENTATION SHOW CLAIMANT WAS PAID FOR ALL COMMISSIONS DUE",
         "" ="OVERPAYMENT TO CLAIMANT",
         "" ="NO JURISDICTION. ALL WORK WAS PERFORMED IN CONNECTICUT",
         "" ="ACCORDING TO EMPLOYER SUBMITTED INFORMATION ALL WAGES HAVE BEEN PAID.",
         "" ="CLAIMANT WAS OVERPAID BY EMPLOYER",
         "" ="CLAIMANT WAS SUSPENDED BY EMPLOYER AND NOT ELIGIBLE TO BE PAID.",
         "" ="CLAIMANT WAS PAID ALL WAGES OWED.",
         "" ="CLAIMANT WAS PAID ALL WAGES PRIOR TO INVESTIGATION.",
         "" ="CLAIMANT WAS PAID ALL WAGES BY EMPLOYER BEFORE CLAIM WAS MADE.",
         "" ="CLAIMANT SUBMITTED NO DOCUMENTATION.",
         "" ="CLAIMANT WAS ARRESTED FOR CRIME AGAINST THE BUSINESS.",
         "" ="SEEKING MONIES FOR PROMISE OF A RAISE AFTER 90 DAYS",
         "" ="EMPLOYER WAS RECOVERING ADVANCED LEAVE. DOCUMENTED.",
         "" ="MONEY WAS LEGALLY TAKEN FROM HOURLY WAGE AND CLAIMANT WAS NOTIFIED BY EMPLOYER.",
         "" ="EMPLOYER PROVIDED DOCUMENTATION OF EMPLOYEE OVERPAYMENT",
         "" ="RECEIVED PTO PAY THAT COVERS OWED WAGES. WORKED IN WASHINGTON DC",
         "" ="CLAIMANT HAS NO DOCUMENTATION OF TIME WORKED.",
         "" ="DOCUMENTATION INDICATES CLAIMANT HAS BEEN PAID.",
         "" ="RECEIVED DOCUMENTATION INDICATES CLAIMANT WAS PAID.",
         "" ="WORK NOT PERFORMED IN VIRGINIA",
         "" ="NO EMPLOYER/EMPLOYEE RELATIONSHIP",
         "" ="CLAIMANT TOOK THE EMPLOYER TO COURT TO SETTLE.",
         "" ="CLAIMANT WAS PAID CORRECTLY DUE TO OUTSTANDING DRAWS"))
```

